SMT-Boosted Security Types and Correctness for Low-Level MPC

### 2025-11-24 toplas26/intro.tex
checkpoint

Together, our contributions \emph{automate} the \emph{compositional} verification of security
properties in $\metaprot$/$\minicat$. In this work, we build on~\cite{skalka-near-ppdp24}
by developing automatic verification approaches that work for arbitrary
finite fields and scale to large protocols. To accomplish this, we develop new
Hoare logics and type systems for compositional verification, and use SMT to automate the checks
that were accomplished by enumeration in prior work. Specifically, we make the
following contributions:
%
% \begin{enumerate}[i.]
% \item In Section \ref{section-smt}, an interpretation of protocols in
%   $\minicat$ as SMT constraints (Theorem \ref{theorem-toeq}) and
%   methodology for verifying correctness properties.
% \item In Section \ref{section-cpj}, a confidentiality type system for
%   $\minicat$ protocols with a soundness property guaranteeing that
%   information is released only through explicit declassifications
%   (Theorem \ref{theorem-cpj}).
% \item In Section \ref{section-ipj}, an integrity type system for
%   $\minicat$ protocols with a soundness property guaranteeing
%   robustness in malicious settings (Theorem \ref{theorem-ipj}).
% \item In Section \ref{section-metalang}, a dependent Hoare type system and
%   algorithm for the $\metaprot$ language that is sound for the
%   confidentiality and integrity type systems (Theorem
%   \ref{theorem-mtj}), with independently verified, compositional $\Pi$
%   types for functions.
% \end{enumerate}
\begin{enumerate}[i.]
\item In Section~\ref{section-smt}, a Hoare logic for Overture protocols, for the compositional verification of both semi-honest and malicious security.
\item In Section~\ref{section-sty}, a novel share type system, enabling security verification for protocols that use secret sharing as a building block; our approach works for any secret sharing scheme and enforces noninterference modulo output.
\item In Section~\ref{section-ipj}, an integrity type system for extending semi-honest security to malicious security.
\item In Section~\ref{section-metalang}, a Hoare logic for Prelude programs with an automated implementation, for the compositional verification of high-level protocol specifications.
\item In Section~\ref{sec:metapr-type-check}, a compositional type-based verification approach for Prelude programs that combines share types with the automated Hoare logic of Section 7.
\end{enumerate}
%
A prior conference version of this work~\cite{skalka-near-esop25}
proposed a whole-program analysis for Prelude programs that ensures
confidentiality, and combined it with an integrity type system to
prove security of Prelude programs. This extended version adds the
compositional Hoare logics presented in Sections 4 and 7, adds the
share type system presented in Section 5, and fully develops the
compositional type-based verification approach in Section 8. In
addition, we develop an implementation capable of automating the
verification of examples given in the prior work, and our extensions
in this work additionally enable the type-based verification of
noninterference modulo output for secret sharing protocols that
leverage Beaver triples (see Section~\ref{section-examples}).

In all cases, our type systems are boosted by verification mechanisms
provided by Satisfiability Modulo Finite Fields \cite{SMFF}. In
Section \ref{section-examples} we develop and discuss extended example
applications of our type analyses to real protocols including
the Goldreich-Micali-Wigderson (GMW), Bendlin-Damgard-Orland-Zakarias
(BDOZ), and Yao's Garbled Circuits (YGC) protocols.

@@

The most closely related work is \cite{skalka-near-ppdp24}, where
foundations for $\minicat/\metaprot$ are introduced including a
verification algorithm. However, this algorithm only works for
$\mathbb{F}_2$-- e.g., binary circuits-- and does not scale due to
exponential complexity, necessitating semi-automated proof
techniques. In contrast, our type systems scale to arbitrary prime
fields and whole program analysis for larger circuits (protocols). So
our work is a significant advancement of $\minicat/\metaprot$.  Our
work is also closely related to Satisfiability Modulo Finite Fields
\cite{SMFF} but only in the sense that we apply their method as we
discuss in Section \ref{section-smt} and subsequently.

>>>>

Together, our contributions \emph{automate} the \emph{compositional} verification of security
properties in $\metaprot$/$\minicat$. In this work, we build on~\cite{skalka-near-ppdp24}
by developing automatic verification approaches that work for arbitrary
finite fields and scale to large protocols. To accomplish this, we develop new
Hoare logics and type systems for compositional verification, and use SMT to automate the checks
that were accomplished by enumeration in prior work. Specifically, we make the
following contributions:
%
% \begin{enumerate}[i.]
% \item In Section \ref{section-smt}, an interpretation of protocols in
%   $\minicat$ as SMT constraints (Theorem \ref{theorem-toeq}) and
%   methodology for verifying correctness properties.
% \item In Section \ref{section-cpj}, a confidentiality type system for
%   $\minicat$ protocols with a soundness property guaranteeing that
%   information is released only through explicit declassifications
%   (Theorem \ref{theorem-cpj}).
% \item In Section \ref{section-ipj}, an integrity type system for
%   $\minicat$ protocols with a soundness property guaranteeing
%   robustness in malicious settings (Theorem \ref{theorem-ipj}).
% \item In Section \ref{section-metalang}, a dependent Hoare type system and
%   algorithm for the $\metaprot$ language that is sound for the
%   confidentiality and integrity type systems (Theorem
%   \ref{theorem-mtj}), with independently verified, compositional $\Pi$
%   types for functions.
% \end{enumerate}
\begin{enumerate}[i.]
\item In Section~\ref{section-smt}, a Hoare logic for Overture protocols, for the compositional verification of both semi-honest and malicious security.
\item In Section~\ref{section-sty}, a novel share type system, enabling security verification for protocols that use secret sharing as a building block; our approach works for any secret sharing scheme and enforces passive security (confidentiality).
\item In Section~\ref{section-ipj}, an integrity type system for extending passive security to malicious security.
\item In Section~\ref{section-metalang}, a Hoare logic for Prelude programs with an automated implementation, for the compositional verification of high-level protocol specifications.
\item In Section~\ref{sec:metapr-type-check}, a compositional type-based verification approach for Prelude programs that combines share types with the automated Hoare logic of Section 7.
\end{enumerate}
%
A prior conference version of this work~\cite{skalka-near-esop25}
proposed a whole-program analysis for Prelude programs that ensures
confidentiality, and combined it with an integrity type system to
prove security of Prelude programs. This extended version adds the
compositional Hoare logics presented in Sections 4 and 7, adds the
share type system presented in Section 5 that enforces a stronger
security property than the type system in~\cite{skalka-near-esop25},
and fully develops the compositional type-based verification approach
in Section 8. In addition, we develop an implementation capable of
automating the verification of examples given in the prior work, and
our extensions in this work additionally enable the type-based
verification of noninterference modulo output for more complex secret
sharing protocols,e.g., that leverage Beaver triples (see
Section~\ref{section-examples}).

In all cases, our type systems are boosted by verification mechanisms
provided by Satisfiability Modulo Finite Fields \cite{SMFF}. In
Section \ref{section-examples} we develop and discuss extended example
applications of our type analyses to real protocols including
the Goldreich-Micali-Wigderson (GMW), Ben-Or-Goldwasser-Widgerson (BGW),
and Bendlin-Damgard-Orland-Zakarias (BDOZ), and report related
performance metrics.

@@

The current work extends previous work \cite{skalka-near-esop25} as
discussed above. Key differences as noted in Figure
\ref{fig-comp-wrap} include a Hoare logic and a more expressive
confidentiality type systems supporting analysis of more complex
programs that enforces passive security in the real/ideal model, vs.~a
weaker gradual release property as in \cite{skalka-near-esop25}.
Furthermore our Hoare logic analysis has been fully implemented in the
current work in an Antlr/Scala/cvc5-based framework.

Besides \cite{skalka-near-esop25} the most closely related work is
\cite{skalka-near-ppdp24}, where foundations for $\minicat/\metaprot$
are introduced including a verification algorithm. However, this
algorithm only works for $\mathbb{F}_2$-- e.g., binary circuits-- and
does not scale due to exponential complexity, necessitating
semi-automated proof techniques. In contrast, our type systems scale
to arbitrary prime fields and whole program analysis for larger
circuits (protocols). So our work is a significant advancement of
$\minicat/\metaprot$.  Our work is also closely related to
Satisfiability Modulo Finite Fields \cite{SMFF} in that we use it as
an enabling technology as discussed in Section \ref{section-smt} and
subsequently.
